---
title: "Standard workflow of scrna_Analaysis"
output: html_document
date: "2025-10-15"
---

#Set working directory
```{r}
setwd("../Analysis/")
```

#Load the libraries
```{r}
library(Seurat)
library(SeuratDisk)
library(tidyverse)

```

#Load the NSCLC 
```{r}
nsclc_raw <- Read10X_h5(filename = "20k_NSCLC_DTC_3p_nextgem_Multiplex_count_raw_feature_bc_matrix.h5")
nsclc_counts <- nsclc_raw$`Gene Expression`
str(nsclc_counts)
```

#Create Seuratobject
```{r}
nsclc_seurat <- CreateSeuratObject(counts = nsclc_counts, project = "NSCLC", min.cells = 3, min.features = 200)
```

# 1. QUALITY CONTROL ----------

Current methods of single cell extraction method atleast takes 6 hours for tissue dissociation, single cell dopamine capture and mrna extraction.

Longer the extraction procedure and vigorous tissue dissociation procedure often results the captured cells in stressed state or killed.

These cells are called low quality cells which need to be removed during QC.

#Quality control metrics

QC metrics - nFeature_RNA / nCount_RNA - [A UMI (Unique Molecular Identifier) - a short, random DNA sequence added to each RNA molecule during scRNA sequencing] / percent.mt

1.Low-quality cells or empty droplets (no cells) often have very few genes (low Feature RNA & COUNT RNA)

2.cell doublets or multipless have high values of Feature RNA & Count RNA

3.Low-quality/dying cells often have high percentage of mitochondrial genes (percent.mt)

#View QC metrics

```{r}
View(nsclc_seurat@meta.data) #Each row reprents a cells - Shows the no of transcripts (or UMI) and genes in one cells 

range(nsclc_seurat$nCount_RNA) # shows minimum and maximum number of transcripts
range(nsclc_seurat$nFeature_RNA) #shows minimum and maximum no of genes
```

# %MT reads
#Computer mitochorrial contribution per cell and add to metadata

```{r}
nsclc_seurat[["percent.MT"]]  <- PercentageFeatureSet(nsclc_seurat, pattern = "^MT-")
View(nsclc_seurat@meta.data)
```

#Visualizing QC metrics 

Violin plot and scatterplot from Seurat package is used.

ncol =3, violin plots will be arranged in 3 columns in the output figure.

geom_smooth(method = "lm") adds a linear regression line (straight line) showing the trend between the two features.

```{r}
#To see the features individually
VlnPlot(nsclc_seurat, features = c("nCount_RNA", "nFeature_RNA", "percent.MT"), ncol = 3) 

#To see the co-relation of features
FeatureScatter(nsclc_seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + geom_smooth(method = "lm")

```

#X axis - no of transcripts, y axis - no fo genes [Each dot represents a cell]

A good qualtiy cells shd follow a straight line. [Meaning it shd have good no of genes and transcripts]

# Artifacts scenario which need to filtered further
1. If the cells are accumulated in lower right corner [150000, 10000], it indicate experiment has captured few no of genes which are sequenced repeatedly - Reason for higher transcripts counts.
2. If the cells are accumulated at left top corner [20000,30000], it indicate experiment has captured high no of genes but not deeply sequenced.

Overall in this case, majoriy of cells follow striaght line, while few loq quaity cells exist which need to be removed.

#Additional qc metrics
1. Ribosomal genes -  Higher Ribosomal genes, Low quality cells
2. Doublet finder - filter doublets in sample

# 2.  FILTERING OF CELSS ---------

Filtering cells contains less than 5 %MT,and genes ranging from 200 to 2500

```{r}
nsclc_seurat <- subset(nsclc_seurat, subset = nFeature_RNA > 200 & nFeature_RNA <2500 & percent.MT < 5 )
```

# 3. NORMALIZING THE DATA -------- Makes cells comparable

Make gene expression values comparable across cells, accounting for differences in sequencing depth or library size. [Some cells may have more total reads/UMIs than others. Without normalization, highly sequenced cells would dominate the data.]

Default normalization method follows log normalization pf (gene expression in each cell/ Total gene expression ) * scaling factor

```{r}
nsclc_seurat <- NormalizeData(nsclc_seurat,normalization.method = "LogNormalize", scale.factor = 10000) 
```
# 4. Filtering highly variable features ----------
Highly variable genes exhibit high cell - cell variantion which highlights the biological signal in sc datasets.

#Identify highly variable genes
```{r}
nsclc_seurat <- FindVariableFeatures(nsclc_seurat, selection.method = "vst", nfeatures = 2000)
```

#Identify and plot top 10 highly variable genes
```{R}
top10 <- head(VariableFeatures(nsclc_seurat), 10)
plot1 <- VariableFeaturePlot(nsclc_seurat)
LabelPoints(plot = plot1, points = top10, repel = TRUE)

```

#The cells in read are 2000 highly variable features

# 5. Scaling ------ Mkaes genes comparabale

To remove unwanted sources of variation due to biological sources like diff cell cyclees for cells  or technical noise due to batch effects.
This need to be scaled, So that highly expressed genes don’t dominate downstream analyses (PCA, clustering, etc.).

#Scaling data across all genes
```{r}
allgenes <- rownames(nsclc_seurat)
nsclc_seurat <- ScaleData(nsclc_seurat, features = allgenes)
str(nsclc_seurat)
```

Under RNA ASSAYS, 3  slots.
1. Count = raw sparse matrix
2. data =  log normalized counts
3.scale.data = scaled data

# 6. Perform Linear dimensionality reduction - PCA

```{r}
nsclc_seurat <- RunPCA(nsclc_seurat, features = VariableFeatures(nsclc_seurat))
```


#Print and visualize PCA results

```{r}
print(nsclc_seurat[["pca"]],dim = 1:5, nfeatures = 5)
DimHeatmap(nsclc_seurat, dims = 1:5, cells = 500, balanced = TRUE)
```
Heat map, X axis indicates cells, y axis indicates PC componenets. Yellow color - High expression, Purple - Low expression

#Determine the dimesnionality of data

Use elbow plot to identify the statiscally signifcant principal components which capture majority of biological signals
```{r}
ElbowPlot(nsclc_seurat)
```

# 7. Clustering 

Step 1 - Find neigbours
Step 2 - Find clusters - 
resolution or granularity of cluster -> Lower the number fewer the cluster and viceversa [check resolution ranges from 0 to 1 or more to see the best clusters ]
Step 3 - Choose bext resolution using Dimplot
```{r}
nsclc_seurat <- FindNeighbors(nsclc_seurat, dims = 1:15)
nsclc_seurat <- FindClusters(nsclc_seurat, resolution = c(0.1,0.3,0.5,0.7,0.9,1))
DimPlot(nsclc_seurat, group.by = "RNA_snn_res.0.1", label = TRUE)
DimPlot(nsclc_seurat, group.by = "RNA_snn_res.0.3", label = TRUE)
DimPlot(nsclc_seurat, group.by = "RNA_snn_res.0.5", label = TRUE)

```
In this case, 0.1 resolution gives good clustering. While in higher resolution similiar cells are shows as distinc clusters. So, going with rsolution 0.1.

#Setting identity of clusters

```{r}
Idents(nsclc_seurat) #Levels indicte the no of clusters for the data
#setting identity using selected resoltuion
Idents(nsclc_seurat) <-  "RNA_snn_res.0.1"
Idents(nsclc_seurat)
```

# Perform Non-linear dimesnionality - TSNEE OR UMAP
After clustering, group cells of simiilar type together at low dimesiionla place -- 

#TO INSTALL UMAp
```{R}
reticulate::py_install(packages = "umap_learn")
```

```{r}
nsclc_seurat <- RunUMAP(nsclc_seurat,dims = 1:15)
DimPlot(nsclc_seurat, reduction = "umap", label = TRUE)
``` 
















